---
title: "hate-visualizations"
format: html
execute:
  warning: false
  echo: false
---

## A Brief Introduction

This document contains a series of visualizations about hate victimization cases in different countries. The data is collected within the framework of our project called **Crossing Borders, Confronting Bias – Researching Culturally Competent Care for Minorities and Marginalized Groups Through the Eyes of Early Career Students**.

```{r}
# Load the libraries

library(ggplot2)
library(tidyverse)
library(readxl)


# Load the data

heatmap <- read.csv("heatmap.csv")
Psy_Pop <- read.csv("psychologists-per-100.csv")
hatedataset <- read_excel("hatedataset.xlsx")
hate_categories <- read_csv("hate_categories_2022.csv")

# Clean the columns

hatedataset <- janitor::clean_names(hatedataset)
hate_categories <- janitor::clean_names(hate_categories)

```

## Top 3 reported hate crime types in each country

This plot shows the three highest reported hate crime types in the countries in relatively broad conceptual categories.

```{r}

# Create the binary variable directly
heatmap <- heatmap %>%
  mutate(Reported = 1)  # all rows represent a reported crime

# Draw Heatmap directly 
ggplot(heatmap, aes(x=Hate_Crime_Category, y=reorder(Country, desc(Country)), fill=factor(Reported))) +
  geom_tile(color="white") +
  scale_fill_manual(values=c("1"="navy"), guide = "none") +  # remove legend
  labs(title="Highest Reported Hate Crime Types by Country",
       x="Hate Crime Category", y="Country") +
  theme_minimal() +
  theme(axis.text.x=element_text(angle=45, hjust=1),
        panel.grid=element_blank(),
        panel.background = element_rect(fill = "white", color=NA),
        plot.background = element_rect(fill = "white", color=NA))
```

## Frequency of Hate Crime Categories in the Dataset

This plot is a continuation of the previous one. It shows what types of hate crimes were most frequently reported as the top three categories in the countries. As we can see, racism and xenophobia are the leading category, followed by LGBTQ+ related hate crime and religious hate crime.

```{r}
# Create a sorted bar plot
ggplot(heatmap, aes(x = reorder(Hate_Crime_Category, table(Hate_Crime_Category)[Hate_Crime_Category]))) +
  geom_bar(fill = "navy") +
  coord_flip() +
  labs(
    title = "Frequency of Hate Crime Categories in the Dataset",
    x = "",
    y = "Consistently Among the Top 3 Hate Crime Types Across Countries"
  ) +
  theme_minimal()
```

## Psychologists needed!

The plot below shows how many psychologists are active per 100k population in each country. India, as expected due to its population, has the lowest ratio of psychologists. Iran is the second to last after Turkey and Germany, which is in urgent need of psychologists. Recently, there have been demands for a reform in the psychology training system of the country.

```{r}

Psy_Pop <- read.csv("psychologists-per-100.csv")

ggplot(Psy_Pop, aes(x = PsyToPop, y = reorder(Country, PsyToPop))) +
  geom_bar(fill = "navy", stat = "identity") +
  labs(
    title = "Number of Psychologists per 100K",
    x = "Psychologists",
    y = "Country"
  ) +
  geom_text(aes(label = PsyToPop), 
            hjust = -0.2,
            size = 3,
            color = "black") +
  theme_minimal()
```

## Urbanization and psychologists

We can expect that with an increase in the ratio of urbanization, which can be an indicator of development, the number of psychologists will increase as well. As you can see in the data, there is not a strong relationship between the number of psychologists and how urbanized the country is (r = 0.20). The low correlation indicates that there are more important factors feeding into the lack of psychologists in different countries, such as the educational system, cultural values, economic status, etc.

```{r}
hatedataset <- hatedataset |> 
  mutate(urban_vs_rural_percent = urban_vs_rural_percent * 100)


ggplot(hatedataset, aes(
  x = psychologist_ratio_per_100k_people,
  y = urban_vs_rural_percent
)) +
  geom_point(color = "navy", size = 3, alpha = 0.8, na.rm = TRUE) +
  geom_smooth(color = "lightblue", method = "lm", se = FALSE) +
  scale_color_discrete(na.translate = FALSE) +
  theme_minimal() +
  labs(title = "Psychologists per 100k regarding Urbanization",
       x = "Psychologists",
       y = "Urbanization %"
       )
```

## Hate victimization trend and psychologists

In this plot, we see the overall distribution of psychologists per 100k based on the victimization trend in each country. What we can learn from the plot is that, first of all, unfortunately, the trend is increasing in most of the countries. Second, a low number of psychologists doesn't always mean an increase in hate victimization. Third, countries that have an increasing trend have more psychologists than other countries in most cases. It reflects the importance of creating groups of experts from different fields to solve this multidimensional problem and not relying only on psychologists. The bad news is that we couldn't find decreasing trends in the countries we collected data from. This can also be an indicator of better data communication and hate vicimization cases being recognized.

```{r}
my_24_colors <- c(
  "#1f78b4", "#33a02c", "#e31a1c", "#ff7f00", "#6a3d9a", "#b15928",
  "#a6cee3", "#b2df8a", "#fb9a99", "#fdbf6f", "#cab2d6", "#ffff99",
  "#8dd3c7", "#bc80bd", "#ccebc5", "#ffed6f", "#f781bf", "#999999",
  "#66c2a5", "#fc8d62", "#a6d854", "#ffd92f", "#e5c494", "#b3b3b3"
)
# Use only filled shapes (21–25). If you have more than 5 countries, you’ll have to recycle or facet.
filled_shapes <- c(21, 22, 23, 24, 25)

ggplot(hatedataset, aes(x = hate_motivated_violence_victimization_trend, 
                        y = psychologist_ratio_per_100k_people)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(width = 0.3, alpha = 0.5, size = 5, 
              aes(fill = country, shape = country), color = "black", stroke = 0.4) +
  scale_fill_manual(values = my_24_colors) +
  scale_shape_manual(values = rep(filled_shapes, length.out = length(my_24_colors))) +
  labs(title = "Psychologists per 100k by Country and Hate Victimization Trend Records",
       x = "Hate Violated Crime / Victimization Trend",
       y = "Psychologists per 100k",
       fill = "Country",
       shape = "Country") +
  theme_minimal()

```

## Official vs. unofficial data sources

A huge challenge for gathering this data was the credibility of the sources. Many countries are not very transparent about the hate crime data. In order to capture the data we wanted, we had to refer to unofficial sources as well and cross-validate them. Here is a plot comparing the unofficial and official source numbers. As illustrated in the plot, most of the data is gathered using both ways.

```{r}

# Official vs Unofficial Data Captured

ggplot(hatedataset, aes(x = data_capture_official)) +
  geom_bar(fill = "navy") +
  labs(
    title = "Official Data Captured",
    x = "",
    y = "Number of Countries"
  ) +
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -0.3) +
  theme_minimal()

ggplot(hatedataset, aes(x = data_capture_unofficial)) +
  geom_bar(fill = "navy") +
  labs(
    title = "Unofficial Data Captured",
    x = "",
    y = "Number of Countries"
  ) +
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -0.3) +
  theme_minimal()
```

## Hate crimes in the countries with more transparent data

```{r}

## Plot Total Hate Crimes
filter(hate_categories,
       subcategory == "Total",
       country != "UK") |> 
  ggplot(aes(x = country, y = value)) +
  geom_bar(fill = "navy", stat = "identity") +
  labs(
    x = "Countries",
    y = "Number of Total Hate Crimes",
    title = "Total Hate Crimes"
  ) +
  geom_text(aes(label = value), vjust = -0.3) +
  theme_minimal()


## Plot Racism
filter(hate_categories,
       subcategory == "Racial",
       country != "UK") |> 
  ggplot(aes(x = country, y = value)) +
  geom_bar(fill = "navy", stat = "identity") +
  labs(
    x = "Countries",
    y = "Number of Hate Crimes based on Race",
    title = "Racial Hate Crimes"
  ) +
  geom_text(aes(label = value), vjust = -0.3) +
    theme_minimal()


## Plot Religion

hate_categories %>%
  filter(subcategory %in% c("Religious", "Anti-Christian", "Anti-Muslim", "Other religion"),
         country != "UK") %>%
  ggplot(aes(x = country, y = value)) +
  geom_bar(stat = "identity", fill = "navy") +
  labs(
    x = "Countries",
    y = "Number of Hate Crimes based on Religion",
    title = "Religious Hate Crimes"
  ) +
    theme_minimal()


## Plot Gender/Sex Related Hate Crime
hate_categories %>%
  filter(subcategory %in% c("Sexual orientation", "Gender-based", "Transgender", "Misogynistic", "Gender/Sexual identity", "Gender-related diversity", "Misandristic"),
         country != "UK") %>%
  ggplot(aes(x = country, y = value)) +
  geom_bar(stat = "identity", fill = "navy") +
  labs(
    x = "Countries",
    y = "Number of Hate Crimes based Gender/Sex",
    title = "Gender/Sex Related Hate Crimes"
  ) +
    theme_minimal()

```

## Country specific plots

Here, we zoom in a bit to identify more nuanced trends and issues of each country. It should be noted that not all the data allowed us to plot based on country. Only a few countries publish regular, clean, and reliable data regarding hate crime, which is a major obstacle.

## Hate-Motivated Crimes by Category in Germany (2022)

```{r}

filter(hate_categories,
       country == "Germany"
) |> 
  ggplot(aes(x = value, y = reorder(subcategory, value))) +
  geom_bar(
    stat = "identity",
    fill = "navy"
    ) +
  geom_text(aes(label = value), 
                 hjust = -0.2,
                 size = 3,
                 color = "black") +
  theme_minimal() +
  labs(
  x = "Number of Hate Motivated Crimes in Germany",
  y = "Categories of Hate Crimes",
  title = "Hate-Motivated Crimes by Category in Germany (2022)" 
)
```

## Hate-Motivated Crimes by Category in Netherlands (2022)

```{r}

filter(hate_categories,
       country == "Netherlands"
) |> 
  ggplot(aes(x = value, y = reorder(subcategory, value))) +
  geom_bar(
    stat = "identity",
    fill = "navy"
  ) +
  geom_text(aes(label = value), 
            hjust = -0.2,
            size = 3,
            color = "black") +
  theme_minimal() +
  labs(
    x = "Number of Hate Motivated Crimes in Netherlands",
    y = "Categories of Hate Crimes",
    title = "Hate-Motivated Crimes by Category in Netherlands (2022)" 
  )
```

## Hate-Motivated Crimes by Category in Canada (2022)

```{r}

filter(hate_categories,
       country == "Canada"
) |> 
  ggplot(aes(x = value, y = reorder(subcategory, value))) +
  geom_bar(
    stat = "identity",
    fill = "navy"
  ) +
  geom_text(aes(label = value), 
            hjust = -0.2,
            size = 3,
            color = "black") +
  theme_minimal() +
  labs(
    x = "Number of Hate Motivated Crimes in Canada",
    y = "Categories of Hate Crimes",
    title = "Hate-Motivated Crimes by Category in Canada (2022)" 
  )
```

## Hate-Motivated Crimes by Category in Italy (2022)

```{r}

filter(hate_categories,
       country == "Italy"
) |> 
  ggplot(aes(x = value, y = reorder(subcategory, value))) +
  geom_bar(
    stat = "identity",
    fill = "navy"
  ) +
  geom_text(aes(label = value), 
            hjust = -0.2,
            size = 3,
            color = "black") +
  theme_minimal() +
  labs(
    x = "Number of Hate Motivated Crimes in Italy",
    y = "Categories of Hate Crimes",
    title = "Hate-Motivated Crimes by Category in Italy (2022)" 
  )
```

## Hate-Motivated Crimes by Category in Norway (2022)

```{r}

filter(hate_categories,
       country == "Norway"
) |> 
  ggplot(aes(x = value, y = reorder(subcategory, value))) +
  geom_bar(
    stat = "identity",
    fill = "navy"
  ) +
  geom_text(aes(label = value), 
            hjust = -0.2,
            size = 3,
            color = "black") +
  theme_minimal() +
  labs(
    x = "Number of Hate Motivated Crimes in Norway",
    y = "Categories of Hate Crimes",
    title = "Hate-Motivated Crimes by Category in Norway (2022)" 
  )
```

## Hate-Motivated Crimes by Category in Austria (2022)

```{r}

filter(hate_categories,
       country == "Austria"
) |> 
  ggplot(aes(x = value, y = reorder(subcategory, value))) +
  geom_bar(
    stat = "identity",
    fill = "navy"
  ) +
  geom_text(aes(label = value), 
            hjust = -0.2,
            size = 3,
            color = "black") +
  theme_minimal() +
  labs(
    x = "Number of Hate Motivated Crimes in Austria",
    y = "Categories of Hate Crimes",
    title = "Hate-Motivated Crimes by Category in Austria (2022)" 
  )
```

## Hate-Motivated Crimes by Category in the UK (2021)

```{r}

filter(hate_categories,
       country == "UK"
) |> 
  ggplot(aes(x = value, y = reorder(subcategory, value))) +
  geom_bar(
    stat = "identity",
    fill = "navy"
  ) +
  geom_text(aes(label = value), 
            hjust = -0.2,
            size = 3,
            color = "black") +
  theme_minimal() +
  labs(
    x = "Number of Hate Motivated Crimes in the UK",
    y = "Categories of Hate Crimes",
    title = "Hate-Motivated Crimes by Category in the UK (2021)" 
  )
```

## Hate-Motivated Crimes by Category in Czech Republic (2022)

```{r}

filter(hate_categories,
       country == "Czech Republic"
) |> 
  ggplot(aes(x = value, y = reorder(subcategory, value))) +
  geom_bar(
    stat = "identity",
    fill = "navy"
  ) +
  geom_text(aes(label = value), 
            hjust = -0.2,
            size = 3,
            color = "black") +
  theme_minimal() +
  labs(
    x = "Number of Hate Motivated Crimes in Czech Republic",
    y = "Categories of Hate Crimes",
    title = "Hate-Motivated Crimes by Category in the Czech Republic (2022)" 
  )
```
